---
title: "SiliconeOil.Microglia.1w"
author: "mphambinh"
date: "2023.11.30"
output: output_format
editor_options: 
  chunk_output_type: console
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

# 0. Shortcuts making your life easier (for Mac)
Option + '-'    : <- 
Cmd + Return    : Run the highlighted command
Cmd + Shift + M : %>% 
 
# 1. Set up your working directory 
Use the command below to set up your working directory. The location you saved your datasets is preferred.
```{r}
rm(list=ls())
# Enter your working directory in to the double bracket ""
knitr::opts_knit$set(root.dir = '/Users/mphambinh/Desktop/SiliconeOil_Project.Microglia/') 
```

# 2. Set up the Seurat Objects
Active the required package
```{r message=FALSE, warning=FALSE}
library(Seurat)
install.packages('Seurat', repos = c('https://satijalab.r-universe.dev', 'https://cloud.r-project.org'))
library(SeuratData)
library(patchwork)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(gridExtra)
BiocManager::install("scater")
library(scater)
library(Seurat)
library(tibble)
library(dplyr)
BiocManager::install("GenomicFeatures")
library(GenomicFeatures)
library(ggplot2)
library(gplots)
library(patchwork)
library(here)
library(cowplot)
#library(monocle3)
library(devtools)
library(Matrix)
library(RColorBrewer)
install.packages("openxlsx")
library(openxlsx)
library(tidyr)
library(plotly)
library(stringr)
library(gridExtra)
library(htmlwidgets)
library(cowplot)
library(htmltools)
install.packages('varSelRF')
library(varSelRF)
library(quantmod)
library(GGally)
library(readr)
library(purrr)
library(fs)
library(sctransform)
install.packages("ggpubr")
library(ggpubr)
library(DoubletFinder)
remotes::install_github('chris-mcginnis-ucsf/DoubletFinder')
library(DoubletFinder)
# # install glmGamPoi
# if (!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")
# BiocManager::install("glmGamPoi")
# # install sctransform from Github
# install.packages("sctransform")
library(sctransform)
library(glmGamPoi)
```
More libraries
```{r}
library(tidyverse)
library(cowplot)
BiocManager::install("edgeR")
devtools::install_url("https://cran.r-project.org/src/contrib/Archive/Matrix.utils/Matrix.utils_0.9.7.tar.gz")
library(Matrix.utils)
library(edgeR)
library(Matrix)
library(reshape2)
library(S4Vectors)
library(SingleCellExperiment)
library(pheatmap)
BiocManager::install("apeglm")
library(apeglm)
library(png)
BiocManager::install("DESeq2")
library(DESeq2)
library(RColorBrewer)
library(data.table)

```

# 3. Load datasets
```{r}
#Load dataset
data_location = '/Users/mphambinh/Desktop/SiliconeOil_Project.Microglia/SO1W'
# change the variable data_location to where you save the Cellranger outputs
dirs <- list.dirs(path = data_location, recursive = F, full.names = F)
# get data location 
for(x in dirs){
  name <- gsub('_filtered_feature_bc_matrix','', x)
  cts <- ReadMtx(mtx = paste0(data_location,"/",x,'/matrix.mtx.gz'),
                 features = paste0(data_location,"/",x,'/features.tsv.gz'),
                 cells = paste0(data_location,"/",x,'/barcodes.tsv.gz'))
  # create seurat objects
  assign(name, CreateSeuratObject(counts = cts
                                  , min.cells = 3 #Only genes/features detected in at least 3 cells will be kept
                                  , min.features=200 #Only cells having at least 200 genes/features will be kept 
  ))}
#Read more : https://www.sciencedirect.com/science/article/pii/S1074761319300731?ref=pdf_download&fr=RR-2&rr=80afafa79d64079a#sec4

#Create universal objects to work with
ctrl <- SO1WCtrl
exp <-  SO1WExp
```
# 4.Create a joint dataset for multiple group 
```{r}
SO1W <- merge(ctrl, exp,
      add.cell.ids = c("ctrl", "exp"),
      project = 'SiliconeOil1Week')

# create a identity column
SO1W$group <- rownames(SO1W@meta.data)

# split sample column
SO1W@meta.data <- separate(SO1W@meta.data, 
                                    col = 'group', 
                                    into = c('Group', 'Barcode'), sep = '_')
View(SO1W@meta.data)
```

# 5.QC
## 5.1. QC for a list of Seurat objects
```{r}
# SO1W.list <- SplitObject(SO1W, split.by = "Group")
# SO1W.list
# #prolif.score = c('pbk', 'birc5a','mki67', 'ube2c', 'top2a', 'tk1', 'aurkb', 'cdkn3', 'cenpf', 'cdk1',"pcna","tubb2b")
# SO1W.list<- lapply(SO1W.list, FUN = function(x) {n
#   x = PercentageFeatureSet(x, pattern = "^mt-", col.name = "percent.mt")
#   x = PercentageFeatureSet(x, pattern = "^Rpl|^Rps", col.name = "percent.ribo")
#   #x = PercentageFeatureSet(x, features = prolif.score, col.name = "prolif.score")
# })
```


## 5.2.1 QC for a merged seurat object
```{r}
# mitopchondria percent
SO1W[["percent.mt"]]<- PercentageFeatureSet(SO1W, pattern = "^mt-")
# ribosome percent
SO1W[["percent.ribo"]] <- PercentageFeatureSet(SO1W, pattern = "^Rpl|^Rps")
# cell cycling score
# need to normalize the dataset first
# SO1W$old.indent <- SO1W$orig.ident
# View(SO1W@meta.data)
# s.genes <- str_to_title(cc.genes$s.genes)
# g2m.genes <- str_to_title(cc.genes$g2m.genes)
# SO1W <- CellCycleScoring(SO1W, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)

```

## 5.2.2 Visualizing QC metrics
```{r}
QC.before.integrate<- VlnPlot(SO1W, 
        features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.ribo"),
        ncol = 2)&
  theme(axis.title.x = element_blank(), axis.text.x = element_text(size = 12, angle = 90, hjust=0.5))
ggplot2::ggsave("QC.before.integrate.jpeg",
                QC.before.integrate,
                width = 12,
                height=10)
SO1W
```
# 6.Filtering
```{r}
SO1w.fil <-  subset(SO1W, subset=
                     nCount_RNA > 1000 & 
                     nFeature_RNA > 250 &
                     nCount_RNA < 50000 &
                     #log10(nFeature_RNA) / log10(nCount_RNA) > 0.76 &
                     percent.mt < 15 
                   )
SO1w.fil
```
# 7.Normalization with SCT
```{r}
# Low resolution
SO1W.fil.normalized = SCTransform(SO1w.fil, vst.flavor = "v2", verbose = FALSE) %>%
    RunPCA(npcs = 30, verbose = FALSE) %>%
    RunUMAP(reduction = "pca", dims = 1:30, verbose = FALSE) %>%
    FindNeighbors(reduction = "pca", dims = 1:30, verbose = FALSE) %>%
    FindClusters(resolution = 0.3, verbose = FALSE)
## Visualize low clustering results
p1 <- DimPlot(SO1W.fil.normalized, label = T, repel = T) + ggtitle("Unsupervised clustering")
p2 <- DimPlot(SO1W.fil.normalized, group.by = "Group") + ggtitle("Clustering by group")
ggplot2::ggsave("clustering.SO1W.fil.normalized.jpeg",
                plot=p1,
                width=10,
                height=8)
ggplot2::ggsave("clustering.by.group.SO1W.fil.normalized.jpeg",
                plot=p2,
                width=10,
                height=8)

# High resolution
ggplot2::ggsave("clustering.SO1W.fil.normalized.jpeg",
                plot=p1,
                width=10,
                height=8)

ggplot2::ggsave("clustering.by.group.SO1W.fil.normalized.jpeg",
                plot=p2,
                width=10,
                height=8)

p1|p2
SO1W.fil.normalized
SO1W.fil.normalized.list <- SplitObject(SO1W.fil.normalized, split.by = "Group")
SO1W.fil.normalized.list
```
# 8. Integration 
```{r}
features = SelectIntegrationFeatures(object.list = SO1W.fil.normalized.list, nfeatures = 3000)
SO1W.fil.normalized.list = PrepSCTIntegration(object.list = SO1W.fil.normalized.list, anchor.features = features)
anchors = FindIntegrationAnchors(object.list = SO1W.fil.normalized.list, normalization.method = "SCT", anchor.features = features)
SO1W.combined = IntegrateData(anchorset = anchors, normalization.method = "SCT")
```
# 9.Clustering
```{r}
SO1W.combined = RunPCA(SO1W.combined, verbose = FALSE)
SO1W.combined = RunUMAP(SO1W.combined, reduction = "pca", dims = 1:35, verbose = FALSE)
SO1W.combined = FindNeighbors(SO1W.combined, reduction = "pca", dims = 1:35)
SO1W.combined = FindClusters(SO1W.combined, resolution = 0.1)
```
Visualization 
```{r}

DimPlot(SO1W.combined, reduction = "umap", label=TRUE, split.by= "Group", pt.size=1)+ NoLegend()
p4<- DimPlot(SO1W.combined, reduction = "umap", label=TRUE, group.by="seurat_clusters")
p3<-DimPlot(SO1W.combined, reduction = "umap", label=TRUE, group.by="Group")
ggplot2::ggsave("clustering.by.group.SO1W.combined.jpeg",
                plot=p3,
                width=10,
                height=8)
ggplot2::ggsave("clustering.SO1W.combined.jpeg",
                plot=p4,
                width=10,
                height=8)

```

# Color
```{r}
mycolor = c("#9ecae1","#c6dbef","#fee8c8","#fcc5c0","#d0d1e6","#fed976","#9e9ac8","#6baed6","#ccece6","#d4b9da","#7bccc4","#ccebc5","#fa9fb5","#feb24c","#fd8d3c","plum","lightskyblue")
mycolor = colorRampPalette(c("lightcoral","wheat1","plum","khaki1","lightskyblue"), bias = 2)(28)
mycolor = c("#E64B35B2","#F39B7FB2","#4DBBD5B2")
gg_color_hue <- function(n) {
  hues = seq(0, 1000, length = n + 1)
  hcl(h = hues, l = 80, c = 40)[1:n]
}
mycolor = c("#ccece6","#99d8c9","#66c2a4","#41ae76","#bfd3e6","#9ebcda","#8c96c6","#8c6bb1","#ccebc5","#a8ddb5","#7bccc4","#4eb3d3","#d0d1e6","#a6bddb","#74a9cf","#3690c0","#fde0dd","#fcc5c0","#fa9fb5","#c994c7","#d4b9da","#c7e9b4","#7fcdbb","#41b6c4","#edf8b1","#ffeda0","#fed976","#feb24c","#fd8d3c","#fee8c8","#fdd49e","#fdbb84","#fc8d59","#fec44f","#fe9929","#c6dbef","#9ecae1","#6baed6","#4292c6","#fcbba1","#fc9272","#9e9ac8","#bcbddc","#d9d9d9","#00AFBB")
```

# 10. Proportion change 
```{r}
SO1W.combined$active.ident <- Idents(SO1W.combined)
View(SO1W.combined@meta.data)
number <- group_by(SO1W.combined@meta.data, Group, active.ident) %>% 
  summarise(count = n()) %>%
  group_by(Group)%>%
  mutate(freq = count/sum(count)*100)
#check
number
#plot cell composition change
  cell_proportion<- ggplot(number, aes(Group, freq, fill = active.ident, width=0.5), label = Cluster) + 
  geom_bar(stat = 'identity', colour="white", alpha=1, linewidth=0.5,position = 'stack') +
  geom_text(aes(label = round(freq,1)), position = position_stack(vjust = 0.5))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  theme_pubclean() +
  theme(legend.position="right")
  cell_proportion
# plot cluster occupancy of each group ( i.e. control and experimental)
number1 <- group_by(SO1W.combined@meta.data, Group, active.ident) %>% 
  summarise(count = n()) %>%
  group_by(Group)%>%
  mutate(perc = round(count/sum(count), 2),
         pos = cumsum(perc) - 0.5*perc)
number1
cluster.occupancy.plot <- ggplot(number1,aes(x = factor(active.ident), fill = factor(Group), y = perc)) + 
geom_col(position = 'fill') +
geom_text(aes(label = paste0(perc*100,"%")), 
            position = position_fill(vjust = 0.5), 
            size = 3) +
labs(fill = 'Group', x = 'Cluster') +
scale_y_continuous(limits = c(0,1))



# save the plots
ggplot2::ggsave("cluster.occupancy.jpeg",
                plot=cluster.occupancy.plot,
                width=10,
                height=8) 
ggplot2::ggsave("cell.proportion.jpeg",
                plot=cell_proportion,
                width=10,
                height=8) 

```


# 11. Annotation 
```{r}
peripheral_macrophage_and_monocyte_marker=c("Ms4a7","Emilin2","Gda", "Hp","Sell")
print(peripheral_macrophage_and_monocyte_marker)
```

Visualize all the clusters we got so far
```{r}
DimPlot(SO1W.combined, split.by = "Group" ,reduction = "umap")
```
Name the clusters"
```{r}
new.cluster.ids <- c("MG0",
                     "MG1",
                     "MG2", 
                     "MG3", 
                     "Photoreceptors",
                     "Proliferating microglia", 
                     "Monocytes/Macrophages")
names(new.cluster.ids) <- levels(SO1W.combined)

SO1W.combined <- RenameIdents(SO1W.combined, new.cluster.ids)
DimPlot(SO1W.combined, reduction = "umap",label.size=3,label = TRUE, pt.size = 0.1)
```

Find markers for : 

1. For one cluster:  
```{r}
cluster0.marker <- FindMarkers(SO1W.combined, ident.1 = 0, min.pct = 0.25 )
head(cluster0.marker, n=10)
VlnPlot(SO1W.combined, features = c(row.names(cluster0.marker)[1], row.names(cluster0.marker)[2], row.names(cluster0.marker)[3]), pt.size=0.01)
```

```{r}
cluster5.marker <- FindMarkers(SO1W.combined, ident.1 = 5, min.pct = 0.25 )
head(cluster5.marker, n=10)
VlnPlot(SO1W.combined, features = c(row.names(cluster5.marker)[1], row.names(cluster5.marker)[2], row.names(cluster5.marker)[3]), pt.size=0.01)
```

2. For all clusters ( each cluster compared to all remaining clusters): 
```{r}
SO.markers <- FindAllMarkers(SO1W.combined, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
```


3. Use reference to define cell-type specific markers.
### Reference
```{r}
#myeloid cells
myeloid_markers = c("Ccr2", "Lyve1","Mgl2","H2-Aa", "F13a1","Mrc1", "Lyz2")
myeloid_featureplot <-FeaturePlot(SO1W.combined, features = myeloid_markers, pt.size = 0.1, label = TRUE)
myeloid_violinplot <-VlnPlot(SO1W.combined, features = myeloid_markers, pt.size = 0.1)
ggsave("myeloid_featureplot.jpeg", myeloid_featureplot)
ggsave("myeloid_violinplot.jpeg", myeloid_violinplot)
#microglia 
microglia_markers= c("Fcrls", "P2ry12", "Cx3cr1", "Trem2", "C1qa")
microglia_featureplot <-FeaturePlot(SO1W.combined, features = microglia_markers, pt.size = 0.1, label = TRUE)
microglia_violinplot <-VlnPlot(SO1W.combined, features = microglia_markers, pt.size = 0.1)
ggsave("microglia_featureplot.jpeg", microglia_featureplot)
ggsave("microglia_violin.jpeg", microglia_violinplot)
#photoreceptors 
photoreceptor_markers= c("Rcvrn", "Pde6g", "Rho")
photoreceptor_featureplot <-FeaturePlot(SO1W.combined, features = photoreceptor_markers, pt.size = 0.1, label = TRUE)
photoreceptor_violin <-VlnPlot(SO1W.combined, features = photoreceptor_markers, pt.size = 0.1)
ggsave("photoreceptor_featureplot.jpeg", photoreceptor_featureplot)
ggsave("photoreceptor_violin.jpeg", photoreceptor_violin)

#resting microglia
resting_markers= c("Lag3", "Egr1", "Ifrd1")
resting_featureplot <-FeaturePlot(SO1W.combined, features = resting_markers, pt.size = 0.1, label = TRUE)
resting_violin <-VlnPlot(SO1W.combined, features = resting_markers, pt.size = 0.1)
ggsave("resting_featureplot.jpeg", resting_featureplot)
ggsave("resting_violin.jpeg", resting_violin)
```

```{r}
proliferating_markers = c("Mki67","Bub1","Pcna")
proliferating_featureplot <- FeaturePlot(SO1W.combined, features = proliferating_markers, pt.size = 0.1, label = TRUE)
proliferating_violinplot <- VlnPlot(SO1W.combined, features = proliferating_markers, pt.size = 0.1)
ggsave("proliferating_featureplot.jpeg", proliferating_featureplot) 
ggsave("proliferating_violinplot.jpeg", proliferating_violinplot) 
```























# DE Analysis
Pseudobulk analysis 
```{r}

SO1W.combined.metadata <- SO1W.combined@meta.data
assay(SO1W.combined.metadata)
cluster_names <- levels(colData(SO1W.combined.metadata)$active.ident)
cluster_names
```




#find marker genes
```{r}
SO1W.combined<- PrepSCTFindMarkers(SO1W.combined)

SO1W.combined.markers <- FindAllMarkers(SO1W.combined, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)

SO1W.combined.markers %>%
  group_by(cluster) %>%
  slice_max(n = 2, order_by = avg_log2FC)
DEG = subset(SO1W.combined.markers, avg_log2FC > 0.5 & p_val_adj < 0.05)

# for one specific subgroup
View(SO1W.combined@meta.data)
DE_group4<-FindMarkers(SO1W.combined,group.by = "",subset.ident=4, ident.1="ctrl", ident.2="exp")
write.xlsx(DEG,file = 'marker.group4.xlsx',rowNames=T)
```



# Testing
```{r}
phagocytosis_related = c("Cd68", "Bin1", "Cd33", "Picalm", "Sorl1" )
spp1_receptor= c("Spp1","Itgb1", "Itgb5", "Ncstn", "Itgav", "Itgb3", "Ccr5", "Ccnd1")
actin_filament_polymerization=c("Kank2", "Twf1")
complement_receptor=c("Cr3")
mechanosensing_related=c("Piezo1","Trpv1","Trpv2", "Trpv4", "Trpm2", "Trpm4", "Trpm7", "Trpc3" )
###
featureplot.microglia<- FeaturePlot(SO1W.combined, features = c(),pt.size=0.1, ncol=2)
vlnplot.microglia<- VlnPlot(SO1W.combined, features = spp1_receptor, ncol=2, pt.size=0.1)
vlnplot.microglia<- VlnPlot(SO1W.combined, features = "Ccl5", ncol=2, pt.size=0.1)
vlnplot.complement <- VlnPlot(SO1W.combined, features=complement_receptor, ncol=2, pt.size = 0.1)
vlnplot.mechanosensing <- VlnPlot(SO1W.combined, features= mechanosensing_related, ncol = 2, pt.size = 0.1)

vlnplot.phagocytosis <- VlnPlot(SO1W.combined, features= phagocytosis_related, ncol = 2, pt.size = 0.1)
###
ggsave("microglia.marker.feature.jpeg", featureplot.microglia)
ggsave("microglia.marker.vln.jpeg",vlnplot.microglia, width=10)
ggsave("microglia.mechanosensing.jpeg", vlnplot.mechanosensing, width =12)
ggsave("microglia.phagocytosis.jpeg", vlnplot.phagocytosis, width =12)
VlnPlot(SO1W.combined, features=peripheral_macrophage_and_monocyte_marker, pt.size=0.1, ncol=2)
```

